---
title: "Homework 3"
author: "Ashwini Varghese"
date: "2022-10-15"
output: github_document
editor_options: 
  chunk_output_type: console
---

## Setup for coding

```{r setup, include = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1



## Problem 2

Let's load, tidy, and wrangling this dataset:

```{r clean}
accel = read_csv("./accel_data.csv") %>% 
    janitor::clean_names() %>% 
    pivot_longer(
      activity_1:activity_1440,
      names_to = "minute",
      names_prefix = "activity_",
      values_to = "activity_count"
    ) %>% 
    mutate(day_type = if_else(day == "Saturday", "weekend",
                              if_else(day == "Sunday", "weekend", "weekday"))) %>%
    mutate(day = as.factor(day)) %>% 
    mutate(minute = as.numeric(minute)) %>% 
    mutate(day_type = as.factor(day_type))

accel
```

Description of dataset:

This dataset has 6 variable (`ncol(accel)`); week, day, and day_id were from the original file untouched besides cleaning up the names using `janitor::clean_names()` but the other 3 (minute, activity_count, and day_type) are new variables created using `pivot_longer` and `mutate`. This dataset has 50,400 observations, found using `nrow(accel)`.





